01 - TreehopperFitness: Quality control

### 2023-07-22 scripts/Publilia-01-qc.Rmd
first full draft of Publilia-01-qc.Rmd script, deleted intermediate checks (the originals of which are all included in the `All_data` .xlsx as sheets)

---
title: '01 - TreehopperFitness: Data prep'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

@@

<!-- This file will not execute by itself. It should be called within the  -->
<!-- runfile to render the final report. You can arrange figures, add explanations, -->
<!-- and do other aesthetic tweaks here. -->

@@

# used 20200319 version
write_csv(Days_observed_dupStatus, here("data/manual_check/Days_observed_dupStatus.csv"))
```

@@

## 2.2 - Create tables to use for manually checking and revising egg status text conflict or missing data
```{r}
# Create table for entries text conflict that need to be resolved
egg_text_conflict_data_v1 <- All_data_v2 %>% # egg_text_conflict: 39 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_text_conflict) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
write_csv(egg_text_conflict_data_v1, here("data/manual_check/egg_text_conflict_data.csv"))

# Create table for entries with missing data that need to be resolved
egg_status_missing_data_v1 <- All_data_v2 %>% # egg_status_missing: 219 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_status_missing, !grepl('NA|M', Female_ID)) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
write_csv(egg_status_missing_data_v1, here("data/manual_check/egg_status_missing_data.csv"))

@@

# 20200326 version was used
write_csv(egg_status_missing_data_eggs_revised, here("data/manual_check/egg_status_missing_data_eggs_revised.csv"))
```

@@

#write_csv(egg_status_missing_data_eggs_revised, "egg_status_missing_data_eggs_revised_20200326.csv")

@@

nymph_status_to_resolve <- nymph_status  %>% # should only include a handful of entries with "?" ambiguous entries for N1 or N2 (because did the revions by hand on 2019/09/26)
  group_by(row_number()) %>%
    filter(any(N1_need_filling, N1_text_conflict, N1_status_missing,
               N2_need_filling, N2_text_conflict, N2_status_missing,
               N5_need_filling, N5_text_conflict, N5_status_missing)) %>%
  ungroup()
# write_csv(nymph_status_to_resolve, "nymph_status_to_resolve.csv") ## manually added this to "nymph_status revisions" sheet in "Watershed_field_data_2019_ALL_local" and manually revised raw data 2019/09/26

@@

# checking number of ambiguous entries that need to be resolved
nymph_text_conflict_data <- All_data_v4 %>% # nymph_text_conflict: 10 on 2019/09/24; 0 on 2019/09/26
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
    filter(any(c(N1_text_conflict, N2_text_conflict, N5_text_conflict))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_text_conflict, N2_text_conflict, N5_text_conflict)
#write_csv(egg_text_conflict_data, "egg_text_conflict_data.csv")

nymph_status_missing_data <- All_data_v4 %>% # nymph_status_missing: 17 on 2019/09/24; 5 on 2019/09/26 (all ok to ignore)
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
  filter(any(c(N1_status_missing, N2_status_missing, N5_status_missing))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_status_missing, N2_status_missing, N5_status_missing)
#write_csv(egg_status_missing_data, "egg_status_missing_data.csv")

@@

## 2.9 - Manually check hatch dates using the "not hatched" notes from field using `check_hatch_date` sheet
Use the "not hatched" notes (from original field notes; typically b/c nymphs present but they are immigrants, not hatched from the focal clutch) to inform first pass of hatch date revisions
Again these changes were specified in the `check_hatch_date` sheet and are applied here without changing the original data file.
```{r}
# Create `check_hatch_date` sheet for checking "not hatched" notes and suggesting manual revisions
check_hatch_date <- All_data_v4  %>%
  filter(grepl("not", Notes, ignore.case = TRUE), grepl("hatch", Notes, ignore.case = TRUE)) %>%
  distinct(location)
#write_csv(check_hatch_date, "check_hatch_date_tmp.csv")
```

@@

>>>>

---
title: '01 - TreehopperFitness: Quality control'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

@@

@@

# used 20200319 version
# write_csv(Days_observed_dupStatus, here("data/manual_check/Days_observed_dupStatus.csv"))
```

@@

## 2.2 - Create tables to use for manually checking and revising egg status text conflict or missing data
```{r}
# Create table for entries text conflict that need to be resolved
egg_text_conflict_data_v1 <- All_data_v2 %>% # egg_text_conflict: 39 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_text_conflict) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
# write_csv(egg_text_conflict_data_v1, here("data/manual_check/egg_text_conflict_data.csv"))

# Create table for entries with missing data that need to be resolved
egg_status_missing_data_v1 <- All_data_v2 %>% # egg_status_missing: 219 on 2019/09/24
  left_join(egg_status_v1, by = c("Date", "location", "Eggs" = "Eggs_status")) %>%
  filter(egg_status_missing, !grepl('NA|M', Female_ID)) %>%
  select(Date, Dorm, Plant, Leaf, Eggs)
# write_csv(egg_status_missing_data_v1, here("data/manual_check/egg_status_missing_data.csv"))

@@

# 20200326 version was used
# write_csv(egg_status_missing_data_eggs_revised, here("data/manual_check/egg_status_missing_data_eggs_revised.csv"))
```

@@

# 20200326 version was used
# write_csv(egg_status_missing_data_eggs_revised, here("data/manual_check/egg_status_missing_data_eggs_revised.csv")
```

@@

nymph_status_to_resolve <- nymph_status  %>% # should only include a handful of entries with "?" ambiguous entries for N1 or N2 (because did the revions by hand on 2019/09/26)
  group_by(row_number()) %>%
    filter(any(N1_need_filling, N1_text_conflict, N1_status_missing,
               N2_need_filling, N2_text_conflict, N2_status_missing,
               N5_need_filling, N5_text_conflict, N5_status_missing)) %>%
  ungroup()
# write_csv(nymph_status_to_resolve, here("data/manual_check/nymph_status_to_resolve.csv") ## manually added this to "nymph_status revisions" sheet in "Watershed_field_data_2019_ALL_local" and manually revised raw data 2019/09/26

@@

# checking number of ambiguous entries that need to be resolved
nymph_text_conflict_data <- All_data_v4 %>% # nymph_text_conflict: 10 on 2019/09/24; 0 on 2019/09/26
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
    filter(any(c(N1_text_conflict, N2_text_conflict, N5_text_conflict))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_text_conflict, N2_text_conflict, N5_text_conflict)
# write_csv(egg_text_conflict_data, "egg_text_conflict_data.csv")

nymph_status_missing_data <- All_data_v4 %>% # nymph_status_missing: 17 on 2019/09/24; 5 on 2019/09/26 (all ok to ignore)
  left_join(N1_status, by = c("Date", "location", "N1")) %>%
  left_join(N2_status, by = c("Date", "location", "N2")) %>%
  left_join(N5_status, by = c("Date", "location", "N5")) %>%
  group_by(row_number()) %>% 
  filter(any(c(N1_status_missing, N2_status_missing, N5_status_missing))) %>%
  ungroup() %>%
  select(Date, Dorm, Plant, Leaf, Female_ID, Eggs, N1, N2, N5, nymph_origin, Notes, N1_status_missing, N2_status_missing, N5_status_missing)
# write_csv(egg_status_missing_data, "egg_status_missing_data.csv")

@@

## 2.9 - Manually check hatch dates using the "not hatched" notes from field using `check_hatch_date` sheet
Use the "not hatched" notes (from original field notes; typically b/c nymphs present but they are immigrants, not hatched from the focal clutch) to inform first pass of hatch date revisions
Again these changes were specified in the `check_hatch_date` sheet and are applied here without changing the original data file.
```{r}
# Create `check_hatch_date` sheet for checking "not hatched" notes and suggesting manual revisions
check_hatch_date <- All_data_v4  %>%
  filter(grepl("not", Notes, ignore.case = TRUE), grepl("hatch", Notes, ignore.case = TRUE)) %>%
  distinct(location)
# write_csv(check_hatch_date, "check_hatch_date_tmp.csv")
```

@@

# 3 - Revise ant data reproducibly
Revised ant data and updated All_data_v5 to All_data_v6
Note that the only changes from All_data_v5 to All_data_v6 were adding columns for ant data and all variables derived from All_data_v5 should be safe to left_join with All_data_v6

## 3.1 - Fill empty entries and resolve text conflicts
As for eggs and nymphs above
```{r}
## find plants where ants were not filled in for all duplicate leaf or plant entries
## for duplicate plant entries
plant_ants_flagged_v1 <- text_conflict_check(All_data_v5, N="Ants_on_plant", perPlant=TRUE) %>%
  select(Date, Dorm, Plant, Ants_on_plant, 
         plant_ants_need_filling = N_need_filling,
         plant_ants_text_conflict = N_text_conflict,
         num_unique_plant_ants = num_unique_N)

plant_ants_fill_source_v1 <- plant_ants_flagged_v1 %>%
  filter(plant_ants_need_filling, !plant_ants_text_conflict, !is.na(Ants_on_plant)) %>%
  select(Date, Dorm, Plant, Ants_on_plant_fill = Ants_on_plant, plant_ants_fill_from_src = plant_ants_need_filling)

All_data_v5_plant_ants_v1 <- All_data_v5 %>%
  left_join(plant_ants_fill_source_v1, by=c("Date", "Dorm", "Plant")) %>%
  mutate(Ants_on_plant_filled = ifelse(is.na(plant_ants_fill_from_src),
                              Ants_on_plant,
                              Ants_on_plant_fill))

# manually check that Ants_on_plant_filled is what you expect before replacing Ants_on_plant
All_data_v5_plant_ants_v1 <- All_data_v5_plant_ants_v1  %>% mutate(Ants_on_plant = Ants_on_plant_filled) %>% select(-Ants_on_plant_filled)

## for duplicate leaf entries
leaf_ants_flagged_v1 <- text_conflict_check(All_data_v5, N="Ants_on_leaf", perPlant=FALSE) %>%
  select(Date, location, Ants_on_leaf,
         leaf_ants_need_filling = N_need_filling,
         leaf_ants_text_conflict = N_text_conflict,
         num_unique_leaf_ants = num_unique_N)

leaf_ants_fill_source_v1 <- leaf_ants_flagged_v1 %>%
  filter(leaf_ants_need_filling, !leaf_ants_text_conflict, !is.na(Ants_on_leaf)) %>%
  select(Date, location, Ants_on_leaf_fill = Ants_on_leaf, leaf_ants_fill_from_src = leaf_ants_need_filling)

All_data_v5_plant_leaf_ants_v1 <- All_data_v5_plant_ants_v1 %>%
  left_join(leaf_ants_fill_source_v1, by=c("Date", "location")) %>%
  mutate(Ants_on_leaf_filled = ifelse(is.na(leaf_ants_fill_from_src),
                                       Ants_on_leaf,
                                       Ants_on_leaf_fill))

# manually check that Ants_on_leaf_filled is what you expect before replacing Ants_on_leaf
All_data_v5_plant_leaf_ants_v2 <- All_data_v5_plant_leaf_ants_v1 %>% mutate(Ants_on_leaf = Ants_on_leaf_filled) %>% select(-Ants_on_leaf_filled)
```

## 3.2 - Check that filled empty entries and text conflicts have been resolved
```{r}
# check that the automatic filling worked and address the text conflicts manually in a separate spreadsheet
plant_ants_flagged_v2 <- text_conflict_check(All_data_v5_plant_leaf_ants_v2,
                                             N="Ants_on_plant", perPlant=TRUE) %>%
  select(Date, Dorm, Plant, Ants_on_plant, 
         plant_ants_need_filling = N_need_filling,
         plant_ants_text_conflict = N_text_conflict,
         num_unique_plant_ants = num_unique_N)

## for duplicate leaf entries
leaf_ants_flagged_v2 <- text_conflict_check(All_data_v5_plant_leaf_ants_v2,
                                            N="Ants_on_leaf", perPlant=FALSE) %>%
  select(Date, location, Ants_on_leaf,
         leaf_ants_need_filling = N_need_filling,
         leaf_ants_text_conflict = N_text_conflict,
         num_unique_leaf_ants = num_unique_N)

ant_text_conflict_v2 <- filter(plant_ants_flagged_v2, plant_ants_text_conflict) %>%
  bind_rows(filter(leaf_ants_flagged_v2, leaf_ants_text_conflict))

# write_csv(ant_text_conflict, "ant_text_conflict.csv") # resolved 2019-10-08
```

## 3.3 - Split ant species into separate columns, check and revise ant counts per species and total
Note: If the total number of ants on plant is less than the sum of the leaves, recalculated ants_P (i.e. ants on plant); ants on stem could be problematic with this, since some groups of ants are assigned to multiple females, but we tried our best to write in the totals for any plant that had such an ambiguous case, so I think I can just omit stem and apex
```{r}
All_data_v5_plant_leaf_ants_split_v1 <- All_data_v5_plant_leaf_ants_v2 %>% 
  mutate(Ants_on_plant_preprocessed = str_to_lower(str_replace_all(str_trim(Ants_on_plant), pattern = " ", replacement = ""))) %>% # remove all whitespace
  mutate(sbP = as.numeric(str_extract(str_extract(Ants_on_plant, "([0-9]+)sb"), "[0-9]+")),
         brbP = as.numeric(str_extract(str_extract(Ants_on_plant, "([0-9]+)brb"), "[0-9]+")),
         brrbP = as.numeric(str_extract(str_extract(Ants_on_plant, "([0-9]+)brrb"), "[0-9]+")),
         srP = as.numeric(str_extract(str_extract(Ants_on_plant, "([0-9]+)sr"), "[0-9]+")),
         lrP = as.numeric(str_extract(str_extract(Ants_on_plant, "([0-9]+)lr"), "[0-9]+"))) %>%
  mutate(sbL = as.numeric(str_extract(str_extract(Ants_on_leaf, "([0-9]+)sb"), "[0-9]+")),
         brbL = as.numeric(str_extract(str_extract(Ants_on_leaf, "([0-9]+)brb"), "[0-9]+")),
         brrbL = as.numeric(str_extract(str_extract(Ants_on_leaf, "([0-9]+)brrb"), "[0-9]+")),
         srL = as.numeric(str_extract(str_extract(Ants_on_leaf, "([0-9]+)sr"), "[0-9]+")),
         lrL = as.numeric(str_extract(str_extract(Ants_on_leaf, "([0-9]+)lr"), "[0-9]+")))# %>%
  #select(-Ants_on_plant, -Ants_on_leaf, -Ants_on_plant_preprocessed)

All_data_v5_plant_leaf_ants_split_v1_revisions <- All_data_v5_plant_leaf_ants_split_v1 %>% 
  group_by(Date, Dorm, Plant, Leaf) %>%
    filter(row_number() == 1) %>%
  ungroup() %>%
  group_by(Date, Dorm, Plant) %>%
    mutate(sbP_revise = ifelse(!grepl("_S|_A", location),
                               ifelse(is.na(sbP) & !all(is.na(sbL)) | sbP < sum(sbL, na.rm = TRUE),
                                      sum(sbL, na.rm = TRUE),
                                      sbP),
                               sbP),
              brbP_revise = ifelse(!grepl("_S|_A", location),
                               ifelse(is.na(brbP) & !all(is.na(brbL)) | brbP < sum(brbL, na.rm = TRUE),
                                      sum(brbL, na.rm = TRUE),
                                      brbP),
                               brbP),
              brrbP_revise = ifelse(!grepl("_S|_A", location),
                               ifelse(is.na(brrbP) & !all(is.na(brrbL)) | brrbP < sum(brrbL, na.rm = TRUE),
                                      sum(brrbL, na.rm = TRUE),
                                      brrbP),
                               brrbP),
              lrP_revise = ifelse(!grepl("_S|_A", location),
                               ifelse(is.na(lrP) & !all(is.na(lrL)) | lrP < sum(lrL, na.rm = TRUE),
                                      sum(lrL, na.rm = TRUE),
                                      lrP),
                               lrP),
              srP_revise = ifelse(!grepl("_S|_A", location),
                               ifelse(is.na(srP) & !all(is.na(srL)) | srP < sum(srL, na.rm = TRUE),
                                      sum(srL, na.rm = TRUE),
                                      srP),
                               srP)) %>%
  ungroup() %>%
  select(Date, location, sbP_revise, brbP_revise, brrbP_revise, srP_revise, lrP_revise) %>%
  filter(!is.na(location))

# Combine revised ant counts for each species with All_data to create `All_data_v6`
All_data_v6 <- All_data_v5_plant_leaf_ants_split_v1 %>% 
  left_join(All_data_v5_plant_leaf_ants_split_v1_revisions, by = c("Date", "location")) %>%
  select(-sbP, -brbP, -brrbP, -srP, -lrP) %>%
  rename(sbP = sbP_revise, brbP = brbP_revise, brrbP = brrbP_revise, srP = srP_revise, lrP = lrP_revise) %>%
  group_by(row_number()) %>%
    mutate(ants_L = sum(sbL, brbL, brrbL, srL, lrL, na.rm = TRUE),
           ants_P = sum(sbP, brbP, brrbP, srP, lrP, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-`row_number()`, -Ants_on_plant, -Ants_on_leaf, -Ants_on_plant_fill, -plant_ants_fill_from_src, -Ants_on_leaf_fill, -leaf_ants_fill_from_src, -Ants_on_plant_preprocessed)
```

# 4 - Final pass at revising clutch status

## 4.1 - Calculate female visit order to help with visualizing egg guarding timelines
```{r}
visit_order_v1 <- All_data_v6 %>%
  filter(!is.na(Female_ID), !grepl('NA|M|_E|_N|N (dup)', Female_ID)) %>%
  select(Date, location, Female_ID) %>%
  arrange(location, Date) %>%
  group_by(location, Female_ID) %>%
    filter(row_number() == 1) %>%
  ungroup() %>%
  group_by(location) %>%
    mutate(visit_order = rank(Date, ties.method = "min")) %>%
  ungroup() %>%
  select(-Date)

# this just helps with plotting timeline, doesn't deal with visit order ties
visitor_rank_v1 <- All_data_v6 %>%
  filter(!is.na(Female_ID), !grepl('NA|M|_E|_N|N (dup)', Female_ID)) %>%
  filter(!grepl("_S", location)) %>%
  select(Date, location, Female_ID) %>%
  arrange(location, Date) %>%
  group_by(location, Female_ID) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  group_by(location) %>%
  mutate(visitor_rank = rank(Date, ties.method = "random")) %>%
  ungroup() %>%
  select(-Date)

Female_location_visit_order_v1 <- All_data_v6 %>%
  filter(!is.na(Female_ID), !grepl('NA|M|_E|_N|N (dup)', Female_ID)) %>%
  filter(!grepl("_S", location)) %>%
  select(Date, location, Female_ID) %>%
  arrange(location, Date) %>%
  left_join(visit_order_v1) %>%
  left_join(visitor_rank_v1)
```

## 4.2 - Identify "clutch fate": when leaves die / go missing / are 100% eaten
```{r}
## I think it would be helpful to identify when leaves die / go missing / are 100% eaten
Eggs_eaten_dates_v1 <- clutch_status_v2 %>%
  select(location, Date, Clutch_status) %>%
  mutate(eggs_eaten = grepl("eaten", Clutch_status, ignore.case = TRUE)) %>%
  filter(eggs_eaten) %>%
  group_by(location) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(location, eggs_eaten_date = Date)

Leaf_missing_dates_v1 <- clutch_status_v2 %>%
  select(location, Date, Clutch_status) %>%
  mutate(leaf_missing_date = grepl("missing", Clutch_status, ignore.case = TRUE)) %>%
  filter(leaf_missing_date) %>%
  group_by(location) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(location, leaf_missing_date = Date)

Leaf_dead_dates_v1 <- clutch_status_v2 %>%
  select(location, Date, Clutch_status) %>%
  mutate(leaf_dead_date = grepl("dead", Clutch_status, ignore.case = TRUE)) %>%
  filter(leaf_dead_date) %>%
  group_by(location) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(location, leaf_dead_date = Date)

egg_fate_dates_v1 <- clutch_nymph_status_v2 %>%
  left_join(Eggs_eaten_dates_v1) %>%
  left_join(Leaf_missing_dates_v1) %>%
  left_join(Leaf_dead_dates_v1) %>%
  group_by(location) %>%
    filter(row_number() == 1) %>%
    mutate(egg_fate = ifelse(any(!is.na(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date))),
                               ifelse(which.min(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date)) == 1, "Eaten/hatched",
                                       ifelse(which.min(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date)) == 2, "Missing",
                                               ifelse(which.min(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date)) == 3, "Dead",
                                                       NA_character_))),NA_character_)) %>%
  mutate(egg_fate_date = if_else(any(!is.na(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date))),
                            min(c(eggs_eaten_date, leaf_missing_date, leaf_dead_date), na.rm = TRUE),
                            as.Date(NA))) %>%
  select(location, egg_fate, egg_fate_date)
```

## 4.3 - Finalizing initiation and hatch dates (plus leaf fate etc) = "clutch status"
```{r}
clutch_status_init_hatch_v1 <- clutch_nymph_status_v2 %>%
  left_join(initiation_dates_v3, by="location") %>%
  left_join(hatch_dates_v3, by="location") %>%
  left_join(egg_fate_dates_v1, by="location") %>%
  mutate(initiation_date_binary = Date == initiation_date,
         hatch_date_binary = Date == hatch_date,
         egg_fate_date_binary = Date == egg_fate_date) %>%
  left_join(Female_location_visit_order_v1, by=c("Date", "location")) %>%
  select(location, Date, Female_ID, visit_order, Clutch_status, egg_num, egg_len, 
         N1, N2, N5, nymph_origin, NOr_ambiguous, initiation_date, initiation_date_binary, hatch_date, hatch_date_binary, visitor_rank,
         egg_fate, egg_fate_date, egg_fate_date_binary)

clutch_status_init_hatch_v1 %>% filter(egg_len > 50)
clutch_status_init_hatch_v1 %>% filter(location == "A61_10") %>% print(n=Inf)

# dates <- as.Date((clutch_status_init_hatch_v1 %>%
#   distinct(Date) %>% arrange(Date))$Date)

clutch_status_last_date_measured_v1 <- clutch_status_init_hatch_v1 %>%
  arrange(location, desc(Date)) %>%
  group_by(location) %>%
    filter(!is.na(egg_len|egg_num)) %>%
    filter(row_number() == 1) %>%
  ungroup()

clutch_status_init_hatch_labels_v1 <- clutch_status_init_hatch_v1 %>%
  mutate(init_label = ifelse(initiation_date_binary,
                             "O",
                             NA_character_),
         hatch_label = ifelse(hatch_date_binary,
                              "H",
                              NA_character_),
         fate_label = ifelse(egg_fate_date_binary,
                              egg_fate,
                              NA_character_)) %>%
  group_by(location, Female_ID) %>%
    arrange(location, Female_ID, desc(Date)) %>%
    mutate(Female_ID_label = ifelse(Date == max(Date),
                                    Female_ID,
                                    NA_character_)) %>%
  ungroup() %>%
  group_by(location) %>%
    mutate(unhatched_days = ifelse(hatch_date_binary, 
                                   paste0(hatch_date-initiation_date, " days"),
                                   NA_character_)) %>%
  ungroup() %>%
  arrange(location, Date)

# exploring the distribution of clutch size metrics
clutch_status_stats_v1 <- clutch_status_init_hatch_labels_v1 %>%
  filter(egg_num > 0 | Clutch_status == "E") %>%
  group_by(location) %>%
    summarize(first_egg_num = first(egg_num),
              min_egg_num = min(egg_num, na.rm = TRUE),
              max_egg_num = max(egg_num, na.rm = TRUE),
              first_egg_len = first(egg_len),
              min_egg_len = min(egg_len, na.rm = TRUE),
              max_egg_len = max(egg_len, na.rm = TRUE),
              initiation_date = first(initiation_date),
              hatch_date = first(hatch_date),
              egg_fate = factor(first(egg_fate)),
              egg_fate_date = first(egg_fate_date),
              any_NOr_ambig_loc = any(NOr_ambiguous)) %>%
  ungroup() %>%
  mutate(unhatched_days = as.integer(hatch_date-initiation_date)) %>%
  left_join(omit_for_abandon_time) %>%
# add number of clutches per plant
  separate(location, c("DormPlant", "Leaf"), "_", remove = FALSE) %>%
  mutate(Dorm = substr(DormPlant, 1,1),
         Plant = substring(DormPlant, 2)) %>%
  group_by(Dorm, Plant) %>%
    mutate(n_clutches_on_plant = n()) %>%
  ungroup() %>%
  select(-Leaf, -DormPlant, -Dorm, -Plant)
```

Note that there is a strong relationship between egg number and clutch length, but the variance increases dramatically at larger numbers of eggs
```{r}
## How tight is the relationship between number of eggs and egg mass length?
ggplot(data = clutch_status_init_hatch_v1, aes(x = egg_len, y = egg_num)) +
  geom_hex() +
  geom_smooth()
```

### [See note] Exporting a list of all clutches with initiation date and hatch date to keep track of revisions
*Note:* Initially I ran this after creating the clutch_status_stats table, but to avoid needing to redo all the calculations/revisions to update clutch_status_stats I moved the data input and revision steps to before I create clutch_status_stats. This would need to be changed if working with the fresh data where "egg_measures_revions" doesn't already exist on the spreadsheet
```{r}
# exporting a list of all clutches with initiation date and hatch date to keep track of revisions
short_unhatched_days_check <- clutch_status_stats_v1 %>% 
  filter(unhatched_days <= 16)

# write_csv(short_unhatched_days_check, "short_unhatched_days_check_tmp.csv")

# *** when is this used for revisions, if ever?
never_hatched_check <- clutch_status_stats_v1 %>% 
  filter(is.na(hatch_date))

# write_csv(never_hatched_check, "never_hatched_check_tmp.csv")

# write_csv(clutch_status_stats_v1, "egg_measures_check_tmp.csv")
```
